### 2018-04-19 Writing/draft.tex
added reproducibility

The resulting full simulation was 4TB @@@ big so is not shared here (though the parameters are).
However, the resulting consensus trees on which the topological differences are calculated are available at @@@.

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
\caption{Effect of character difference on recovering the ``normal'' topology. The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
\label{Fig:Tr_results_best}
\end{figure}

@@

\subsubsection{Number of characters}
This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively - see supplementary materials 3) with a significant difference only between 100 and 1000 and 350 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
%MB: The above sentence has too many "for"s. It's too long and needs to be split to be intelligible.
%TG: fix attempt
The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.

@@

\subsection{Effect of character differences on topology}
As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
When characters are correlated, one could expect the matrices to convey a really strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary materials @@@).

@@

Although we measured a significant difference between ``small'' and larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distribution that is always really high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
For the number of characters, this suggests that the overall character difference metric is a good proxy for character correlation as it is independent of the number of characters analysed.
Similarly, using the a Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e. we corrected for the expected difference when comparing two random trees with the same number of taxa).

\subsection{Effect of character differences on the inference method}
When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the maximum parsimony trees in terms of clade conservation but none in terms of taxa placement (both using a Wilcoxon test and the Bhattacharrya Coefficient; Table \ref{Tab_pooledsmethods_test}).
The difference in the ability of each method to recover the ``correct'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have raised issues with these investigations \citealt{spencerefficacy2013,goloboff2017weighted}).
In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over Parsimony, however, our protocol for selecting matrices \citep[i.e. those with in a $CI<0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour Parsimony analysis.
It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.

%MB: I'm not sure I'm convinced by this paragraph. Perhaps we should have a chat about it.
%TG: I've rephrased the parargaph below. This is the idea (a bit arms waving, I agree): for a given topology, when there is high homoplasy, parsimony will give the tree a high score (may changes per branches - even though branch length is not a parsimony parameter) and probably discard the topology. For the same topologyn and high homoplasy, Bayesian will stretch the branches (accounting for the many changes per branches) which will decrease the likelihood score but less fast than MP (because branch length is only *one* of the parameters accounting for the likelihood score). Also, in Bayesian, an decrease in likelihood score does not necessary mean that the tree will be discarded.
The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate score differently:
(1) in parsimony, the topology will probably be given a bad optimality score (on that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
The tree search will thus likely result in a topology island that will not contain the given topology $T$.
(2) in Bayesian inference, the topology will also be given a bad optimality score (i.e. low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
The rate and the branch length being two parameters among others, the optimality score (the likelihood) will change less drastically than for using parsimony.
Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.

@@

\section{Data availability, repeatability and reproducibility}
The consensus trees are available on figshare at \url{@@@}.
The simulations are fully replicable following the explanations at \url{https://github.com/TGuillerme/CharactersCorrelation}.
The post-simulation analysis, tables and figures (reported in this manuscript) are fully reproducible see (\url{https://github.com/TGuillerme/CharactersCorrelation}).

>>>>

The resulting full simulation was 3.5TB big so is not shared here (though the parameters are).
However, the resulting consensus trees on which the topological differences are calculated are available at \url{https://figshare.com/s/7a8fde8eaa39a3d3cf56}.

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf}
\caption{Effect of character difference on recovering the ``normal'' topology. The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
\label{Fig:Tr_results_best}
\end{figure}

@@

\subsubsection{Number of characters}
This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively - see supplementary materials 3) with a significant difference only between 100 and 1000 and 350 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to really similar distributions.

@@

\subsection{Effect of character differences on topology}
As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary materials @@@).

@@

Although we measured a significant difference between ``small'' and larger matrices (both in terms of number of taxa and characters; Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}), these differences have to be contrasted with the probability of overlap between the results distributions that are always really high between every matrices sizes (Bhattacharrya Coefficients $>0.965$ for both the different number of characters and taxa).
This suggest that the effect of character correlation on recovering the right topology is independent of the size of the matrix when pooling the data.
For the number of characters, this suggests that the overall character difference metric is a good proxy for character correlation as it is independent of the number of characters analysed.
Similarly, using the a Normalised Tree Similarity metric ($NTS$) accounts for the fact that topological difference is affected by the sheer number of taxa considered (i.e. we corrected for the expected difference when comparing two random trees with the same number of taxa).

\subsection{Effect of character differences on the inference method}
When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the maximum parsimony trees in terms of clade conservation but none in terms of taxa placement (both using a Wilcoxon test and the Bhattacharrya Coefficient; Table \ref{Tab_pooledsmethods_test}).
The difference in the ability of each method to recover the ``correct'' topology has been heavily discussed in the last five years with some indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; although some critics have raised issues with these investigations \citealt{spencerefficacy2013,goloboff2017weighted}).
In this study, it is possible that our simulation protocol for generating the characters (favouring slightly more M$k$-based characters rather than HKY ones) could slightly favour Bayesian inference over maximum parsimony, however, our protocol for selecting matrices \citep[i.e. those with in a $CI<0.26$ in a quick preliminary parsimony search;][]{OReilly20160081} could also favour maximum parsimony analysis.
It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.

The differences in performance of the two methods observed here could be due to the inherent mechanisms of each method.
For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate score differently:
(1) in parsimony, the topology will probably be given a bad optimality score (on that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the tree.
The tree search will thus likely result in a topology island that will not contain the given topology $T$.
(2) in Bayesian inference, the topology will also be given a bad optimality score (i.e. low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
The rate and the branch length being two parameters among others, the optimality score (the likelihood) will change less drastically than for using parsimony.
Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
This difference in both mechanisms could explain why, on average, Bayesian Inference seems better to recover the ``normal'' topology than maximum parsimony.

@@

\section{Data availability, repeatability and reproducibility}
The consensus trees are available on figshare at \url{https://figshare.com/s/7a8fde8eaa39a3d3cf56}.
The simulations are fully replicable following the explanations at \url{https://github.com/TGuillerme/CharactersCorrelation}.
The post-simulation analysis, tables and figures (reported in this manuscript) are fully reproducible see (\url{https://github.com/TGuillerme/CharactersCorrelation}).
